---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Assignment Week 2

## 1) Loading the data

```{r}
stepdata <- read.csv("Assignment_Week2.csv", header = TRUE)
head(stepdata)
```

## 2) Calculate total number of steps taken each day

loading all the needed packages
```{r, echo = FALSE}
library(magrittr)
library(dplyr)
library(ggplot2)
```

```{r}
databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% 
  summarize(tsteps= sum(steps)) %>%na.omit()
hist(databydate$tsteps, xlab = "Total daily Steps",
     main="Histogram of Total Steps by day", breaks = 20)
```

## 3) Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(databydate$tsteps)

median(databydate$tsteps)
```

## 4) Time series plot

```{r}
databyinterval <- stepdata%>% select(interval, steps) %>% 
  na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps)) 

ggplot(databyinterval, aes(x=interval, y=tsteps))+ geom_line()
```

## 5) The 5-minute interval that, on average, contains the maximum number of steps

```{r}
databyinterval[which(databyinterval$tsteps== max(databyinterval$tsteps)),]
```

# Dealing with missing values

## 6) Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

### Generate listing of NA's

```{r}
missingVals <- sum(is.na(data))

missingVals
```

### Devise a strategy for filling in all of the missing values in the dataset

```{r}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```

## 7) Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
```

### Summary(histogram) of new data : mean & median

```{r}
summary(FullSummedDataByDay)

hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency",
     main = "Total Daily Steps", breaks = 20)
```

### Compare the mean and median of old and new data

```{r}
oldmean <- mean(databydate$tsteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)

oldmean
newmean

oldmedian <- median(databydate$tsteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)

oldmedian
newmedian
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )

meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")
```

## 8) Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
  facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
  ggtitle("Comparison of Average Number of Steps in Each Interval")
```
